Adaptive Waveform Inversion for Transmitted Wavefields

### 2024-10-16 repro/papers/awi/paper.tex
close

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution
solution if $f \in \cal{E}'(\bR)$ (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (for example, \cite{Tay:81}, Theorem VI.2.1)
implies that if $f \in \cal{E}'(\bR)$ and $(\bx_s,\bx_r) \in X_{sr}$,
then the trace $p(\bx_r, \cdot;\bx_s)$ on $\{(\bx_r,t): t \in \bR\}$ is well-defined,
as a distribution vanishing for sufficiently large
negative $t$. Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
  \rightarrow X_{sr} \times {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

In this setting, acoustic FWI means: given the set of
source-receiver position pairs $X_{sr}$, source wavelet
$f$ subject to constraints F1 and F2,  density $\rho$, and data traces
$d:  X_{sr} \rightarrow \cal{D}(\bR)$, find $\kappa$ (or equivalently
so that $F[\kappa] \approx d$ in some sense. If
$d$ and $F[\kappa]$ lie in a normed subspace of $\cal{D}(\bR)$ for each
$(\bx_s,\bx_r) \in X_{sr}$, then the
norm provides a straightforward candidate for this sense of
approximation.

@@

\subsection{Hadamard's Decomposition}
The Hadamard decomposition of the acoustic Green's function provides a
basis for a norm-based formulation of FWI \cite[]{Friedlander:75,Qian:JCP24}. As presented by
\cite{Friedlander:75}, for instance, this construction applies to
second-order wave equations, but the system \ref{eqn:awe} is
equivalent fo a second-order problem: differentiate the first equation
with respect to $t$, eliminate $\bv$ via the second equation, and
rearrange to obtain
\begin{equation}
  \label{eqn:awe2}
  \frac{\rho}{\kappa}\frac{\partial^2 p}{\partial t^2} - \nabla^2 p -
  \nabla \log \rho \cdot \nabla p = \rho \frac{\partial f}{\partial t}
  \delta(\bx-\bx_s), \nonumber \\
  p=0, \, t \ll 0.
\end{equation}
If $(p,\bv)$ is a distribution solutions of the system
\ref{eqn:awe} then $p$ is a solution of equation
\ref{eqn:awe2}, as follows by differentiating the first equation of
\ref{eqn:awe} with respect to $t$ and eliminating $\bv$ via the second
equation. Conversely, if $p$ is a solution of the second-order initial
value problem\ref{eqn:awe2}, it follows from the Hadamard
decomposition detailed below and condition F1
that 
\[
  \bv = \beta \int_{-\infty}^t p
\]
has an
integrable singularity at $\bx=\bx_s$, hence 
defines a distribution, and that $(p,\bv)$ solves the system \ref{eqn:awe}.

Hadamard's construction is local, in both space-time and in the set of
acoustic models. A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray of
geometric acoustics (a geodesic for the Riemannian metric associated
to the equation \ref{eqn:awe2}) (\cite{Friedlander:75},
p. 16). Hadamard's decomposition holds in a space-time cylinder
$\Omega \times \bR$,  and $\Omega\subset \bR^3$ 
and $M \subset (C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3))^2$ are bounded open sets satisfying 
\begin{itemize}
  \item[G1. ] for every $(\bx_s,\bx_r) \in X_{sr}$, $\bx_s, \bx_r \in \Omega$;
  \item[G2. ] for every source point $\bx_s\in X_{s}$ and $\kappa \in M$, $\Omega$ is a
    normal neighborhood of $\bx_s$.
  \item[G3. ] for every source point $\bx_s \in X_{s};$
    $\kappa \in M$, ray ${\bf X}:\bR^+\rightarrow \bR^3$
    with ${\bf X}(0)=\bx_s$ and any $t>0$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
    or $\{{\bf X}(s): t < s < \infty\} \subset \bR^3 \setminus \Omega$.
  \end{itemize}
Condition G3 (the ``no return'' condition) means that once a ray
leaves $\Omega$, it does not return to $\Omega$. Conditions G1 - G3
hold if $\kappa$ is constant, and for small perturbations and some
large perturbations of constant $\kappa$: these are open conditions in
the topology of $M$.

The arguments to follow depend an additional property of solutions of
the wave equation \ref{eqn:awe2}:
\begin{itemize}
\item[G4. ] For any $n > 0$, $f \in {\cal E}'(\bR)$, there exists $K, T$ and $\delta>0$ so
  that for any 4-multiindex ${\bf m}$ of order $\le n$, $(\bx_s,\bx_r) \in
 X_{sr}, t \in \bR$,
  and any $\kappa \in M$, the solution $p$ of \ref{eqn:awe2} is smooth
  for $t>T$ and satisfies
  \[
    |D^{\bf m}p(\bx_r,t;\bx_s)| \le K e^{-\delta t}.
  \]
\end{itemize}
Condition G4 is a special case of {\em local energy decay}, a much
studied topic. Properties which imply G4 are in turn implied by the
{\em non-trapping} hypothesis on the model $\kappa \in M$: all rays of
geometric asymptotics escape to infinity (\cite{Hristova:09}, Theorem
2, \cite{EgorovShubin}, Theorem 2.104, p. 192).

Under conditions G1 - G4, the construction described by
\cite{Friedlander:75} produces for each $m \in M$ functions $\tau[\kappa]$ (travel time), $a[\kappa]$
(geometric amplitude) on $X_s \times \Omega$ and $b[\kappa]$
(remainder kernel) on $X_s \times \Omega \times \bR$, satisfying
\begin{itemize}
\item[H1. ] $\tau[\kappa], a[\kappa],$ and $b[\kappa]$ are smooth off the diagonal,
that is, $\tau[\kappa](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus \{\bx_s\})$
and similarly for $a[\kappa],b[\kappa]$;
\item[H2. ] there exists $A>0$ so that for any $\kappa \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,
  $|\log a[\kappa](\bx_s,\bx_r)| < A$;
\item[H3. ] there exist $B \ge 0,\delta > 0$ so that for
$(\bx_s,\bx_r) \in X_{sr}$, and $\kappa \in
M$,
\begin{eqnarray}
  |b[\kappa](\bx_r,t;\bx_r,s)| &\le& B e^{-\delta t} \nonumber \\
  |\partial_t b[\kappa](\bx_r,t;\bx_r,s)| &\le& B e^{-\delta t}
  \label{eqn:best}
\end{eqnarray}
\end{itemize}
(The exponential decay estimate in item H3 follows from hypothesis G4.)

The Hadamard
decomposition of the Green's function $G[\kappa]$, i.e. the solution of the
equation \ref{eqn:awe2} for $f = \beta H$, is::
\begin{equation}
  \label{eqn:green}
  G[\kappa](\bx_s,\bx,t) = a[\kappa](\bx_s,\bx)\delta(t - \tau[\kappa](\bx_s,\bx)) +
  b[\kappa](\bx_s,\bx,t-\tau[\kappa])H(t - \tau[\kappa](\bx_s,\bx))
\end{equation}
%Under conditions G1 - G3 above, this decomposition is the restriction of
%the Greeen's function to $\Omega \times (-\infty,t_h]$. This fact
%implies a decomposition of $F[\kappa]$, provided that the convolution $G*f$
%restricted to $[0,t_d]$ is determined by the restriction of $G$ to
%$\Omega \times [0,t_h]$. That is the case if $t_h > t_d + |\mbox{supp
%|f/2$ (the factor of $1/2$ because $f$ is odd).
%Making this assumption (which is weaker than assumptions on $t_h$ to be made below),
%\[
%   F[\kappa](\bx_s,\bx_r)(t) =
%   \phi(t)a(\bx_s,\bx_r)\rho_s\frac{df}{dt}(t-\tau[\kappa](\bx_s,\bx_r))
% \]
%\begin{equation}
%  \label{eqn:fwdgreen}
% + \phi(t)\rho_s\int
%  \,ds \, \frac{df}{dt}(t-s) b[\kappa](\bx_s,\bx_r,s-\tau[\kappa](\bx_s,\bx_r))H(s-\tau[\kappa](\bx_s,\bx_%r))
%\end{equation}

\subsection{FWI via Least Squares}
Suppose that the source wavelet $f$ satisfies
\begin{itemize}
\item[F1. ] $f \in H^1_0(\bR)$, $\mbox{supp } \subset [-t_w,t_w]$;
\item[F2. ] $f$ is odd, $f(t) = -f(-t)$.
\end{itemize}
Then the Hadamard decomposition 
(equation \ref{eqn:green}) together with local
energy decay (assumption G4) and Young's inequality  implies that
\begin{equation}
  \label{eqn:fwdgreen}
  F[\kappa] = G[\kappa] * \rho \partial_t f \in L^p(X_{sr} \times \bR)
\end{equation}
for any $p \ge 1$, and that
the corresponding estimates are uniform in $\kappa \in M$:
there exists $C>0$ so that for $\kappa\in M$, $p \ge 1$,
\begin{equation}
  \label{eqn:fwdnorm}
  \|F[\kappa]\|_{L^p(X_{sr} \times \bR)} \le C \left\|\frac{df}{dt}\right\|_{L^p(\bR)}.
\end{equation}
%\[
%  \| \cdot \|_d = \| \cdot \|_{L^2(X_{sr} \times \bR)}
%\]

This observation justfies posing FWI as a least-squares problem: given $d \in
L^2(X_{sr}\times \bR)$, $f \in H^1_0(\bR)$, and $\rho_s>0$, find $\kappa$ to minimize
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[\kappa]-d\|_{L^2(X_{sr} \times \bR)},
\end{equation}
%in which we have introduced the abbreviation
%\[
%  \| \cdot \|_d = \| \cdot \|_{L^2(X_{sr} \times \bR)}
%\]
Practical versions of least-squares FWI include various types of regularization and
imposition of constraints on $m$, see the references cited in the
Introduction.
%As mentioned there, local iterative optimization methods
%are the only practical approach to minimizing $J_{\rm FWI}$ due to the
%computational cost of approximating the forward map $F$ and related
%operators, but these tend to stagnate at highly sub-optimal
%(``cycle-skipped'') estimates of $m$ unless the initial estimate is of
%quite high quality, often available only at considerable cost in data
%collection and processing.

Note that in the representation \ref{eqn:fwdgreen}, only the
derivative $\partial_tf$ appears. For convenience, introduce
\begin{equation}
  \label{eqn:wavelet}
  w = \rho \partial_t f
\end{equation}
The properties F1 and F2 assumed for $f$ are equivalent to these
properties of $w$:
\begin{itemize}
\item[W1. ] $w \in L^2_0(\bR)$, $\mbox{supp }w \subset [-t_w,t_w]$;
\item[W2. ] $w$ is even;
\item[W3. ] $\int \, w = 0$.
\end{itemize}
I will also
refer to $w$ as ``the wavelet'', and view $F$ as parametrized by $w$.

Introduce asymptotics in this setting by choosing $w = \wl$
from a family $\{\wl:\}$ indexed by the dimensionless parameter
$0 <\lambda \le 1\}$, and defined by
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \lambda^{-1/2}w_1(t/\lambda).
\end{equation}
The  ``mother wavelet'' $w_1$ is square-integrable of compact support
even, and mean-zero. So that property W1 holds for the entire family, 
\begin{equation}
  \label{eqn:wdiam}
  \mbox{supp }w_1 \subset [-t_w,t_w].
\end{equation}
All wavelets $\wl$ in this family
inherit these properties from $w_1$, along with its dimensions
(energy/time$^2$). It follows from these assumptions that the Fourier transform $\fwl$ is
real-valued, even, of class $H^1(\bR)$. The corresponding source waveform $f_{\lambda}$ is
given by
\begin{equation}
  \label{eqn:flambda}
  f_{\lambda}(t) = \int_{-\infty}^t\wl =
  \lambda^{1/2}f_1\left(\frac{t}{\lambda}\right).
\end{equation}
Note that for $p \ge 1$,
\begin{equation}
  \label{eqn:lambdanorms}
  \|\wl\|_{L^p(\bR)} = \lambda ^{\frac{1}{p}-\frac{1}{2}}\|w_1\|_{L^p(\bR)},\,\|f_{\lambda}\|_{L^p(\bR)}
  = \lambda ^{\frac{1}{p}+\frac{1}{2} \|f_1\|_{L^p(\bR)}}.
\end{equation}

The parameter $\lambda$ is connected to several frequency-related
characteristics of $\wl$. Suppose that $f \in L^2(\bR)$ is also square-integrable after
multiplication by $t$ (true if  for instance $f$ has compact
support). The {\em pulse width} (or {\em RMS width} ) of $f$ is
\begin{equation}
  \label{eqn:pw}
  l(f) = \frac{\|tf\|}{\|f\|}.
\end{equation}
The pulse width of $\wl$ is proportional to $\lambda$:
\begin{equation}
  \label{eqn:wpw}
  l(\wl) = \left(\frac{\int dt (t \wl(t))^2}{\int dt (\wl(t))^2}
  \right)^{1/2} = \lambda l(w_1).
\end{equation}
The {\em RMS frequency} of $\wl$ is
\begin{equation}
  \label{eqn:freq}
  k(\wl) = \left(\frac{\int d\omega (\omega \fwl(\omega))^2}{\int dt (w(t))^2} \right)^{1/2} = \lambda^{-1}k(w_1).
\end{equation}
Thus the {\em RMS wavelength} $1/k(\wl)$ is proportional to $\lambda$.

The modeling operator with wavelet $\wl$ is denoted $\Fl$. From the
decomposition \ref{eqn:fwdgreen}, $\Fl$ is the sum of two terms:
\begin{equation}
  \label{eqn:Flamdecomp}
  \Fl = \Fla + \Flb.
\end{equation}
where
\begin{equation}
  \label{eqn:Fl0}
  \Fla[\kappa](\bx_r,t;\bx_s) =
  a[\kappa](\bx_s,\bx_r)\wl(t-\tau[\kappa](\bx_s,\bx_r))
\end{equation}
and
\[
\Flb[\kappa] = \rho_s\int
   \,ds \, \frac{df_{\lambda}}{dt}(t-s) b[\kappa](\bx_s,\bx_r,s-\tau[\kappa](\bx_s,\bx_r))H(s-\tau[\kappa](\bx_s,\bx_r))
\]
\[
  =\rho_s
  f_{\lambda}(t-\tau(\bx_s,\bx_r))b[\kappa](\bx_s,\bx_r,0)
\]
\begin{equation}
  \label{eqn:Fl1}
    -
    \rho_s\int
   \,ds \, f_{\lambda}(t-s) (\partial_t
   b[\kappa](\bx_s,\bx_r,s-\tau[\kappa](\bx_s,\bx_r)))H(s-\tau[\kappa](\bx_s,\bx_r))
\end{equation}
From hypothesis G4, the bound \ref{eqn:lambdanorms} , and Young's inequality, the $\|\Fl1[\kappa]\|_{ L^p(X_{sr} \times \bR)}$ is bounded by a multiple of
$\lambda^{\frac{1}{p}+\frac{1}{2}}$, uniformly over $\kappa \in M$.
I will write ``$+ \Ol$'' to mean an  error uniformly bounded in $L^2$ by
a multiple of $\lambda$ over $m \in M$. Thus
\begin{equation}
  \label{eqn:fwdapprox}
  \Fl[\kappa](\bx_s,\bx_r) = \Fla[\kappa] +   \Ol
\end{equation}

@@

For such extended sources,
\[
  \bar{F}[\kappa,u*w] = u*F[\kappa]
\]
Introduce the notation $S[\kappa]$ for the operator of convolution
with $F[\kappa]$, that is,
\begin{equation}
  \label{eqn:sdef}
  S[\kappa]u \equiv u*F[\kappa]
\end{equation}
Then $S[\kappa]$ is bounded on $L^2(X_{sr} \times \bR)$, uniformly in
$\kappa \in M$, thanks to
the bound \ref{eqn:fwdnorm} for $p=1$. $S$ also represents an
extension of $F$, in the sense that with the choice
$u(\bx_r,t;\bx_s)=\delta(t)$, the output of $F$ is recovered:
\begin{equation}
  \label{eqn:sconsist}
  S[\kappa]\delta = F[\kappa].
\end{equation}

%Since $S[\kappa]$ involves solution of the wave equation with point source
%wavelet $u*w$, application of the Hadamard decomposition \ref{eqn:green} requires a furnther
%constraint on ray geometry, expressed by condition G3. Specifically,
%\begin{itemize}
%\item[G4. ]$t_h > t_d + t_w + t_u$.
%\end{itemize}
%Then \ref{eqn:green} implies that
%\begin{equation}
%  \label{eqn:sop}
%  S[\kappa]u = \phi w* a[\kappa] T[\kappa] ((\delta + b[\kappa]H)*u)
%\end{equation}
%in which $T[\kappa]$ is the time translation operator by $\tau[\kappa]$:
%\begin{equation}
%  \label{eqn:Tdef}
%  T[\kappa]u(\bx_r,t;\bx_s) = u(\bx_r,t-\tau[\kappa](\bx_s,\bx_r);\bx_s),
%\end{equation}
%Similarly,
%\begin{equation}
 % \label{eqn:fwdgreenop}
%  F[\kappa] = \phi a[\kappa] T[\kappa]((\delta + b[\kappa]H)*w) 
%\end{equation}

@@

\cite{Warner:16} assume (implicitly) that there is always a unique
solution of the extended inversion problem, formulated in terms of
$S$: that is, that
for any $\kappa$ and any data $d$, there is an
adaptive filter $u[\kappa;d]$ for which
$S[\kappa]u[\kappa;d]= d$. If $d$ is noise-free, that is, $d=F[\kappa]$, then the consistency relation \ref{eqn:sconsist}
suggests that $u[\kappa;d](\bx_r,t;\bx_s) = \delta(t)$. This ideal adaptive filter is in the null
space of multiplication by $t$, that is, the operator
$u(\bx_r,t;\bx_s) \mapsto t u(\bx_r,t;\bx_s)$. This
observation motivates the definition of the AWI objective function
(\cite{Warner:16}, equation XXXX)::
\begin{equation}
  \label{eqn:awidef}
  J_{\rm AWI}[m;d] = \sum_{\bx_s,\bx_r} \left(\frac{\int dt |tu[\kappa;d](\bx_r,t;\bx_s)|^2}{\int dt |u[\kappa;d](\bx_r,t;\bx_s)|^2}\right) 
\end{equation}
in which $u[\kappa;d]$ is as defined above, that is, the solution of $S[\kappa]u[\kappa;d]
= d$. \cite{Warner:16} also formulate
AWI in terms of the inverse filter, that is, $u$ for which
$\phi u*d = F[\kappa]$, and in fact recommend that definition. The present
discussion uses only the forward filter.
Adaptive waveform inversion consists in iterative minimization of
$J_{\rm AWI}$: choose an initial estimate $m_0=(\kappa_0,\beta, w)$ of the
model, then update $m$ (that is, $\kappa$) repeatedly using a gradient
descent method. \cite{Warner:16} describes the computation of $J_{\rm AWI}$
and its gradient in detail, and their use in a quasi-Newton
optimization algorithm.

>>>>

The parameter to be determined in the inverse problem studied here is
the bulk modulus $\kappa$. It is presumed is bounded and smooth:
\begin{equation}
  \label{eqn:coeff}
  \log \kappa \in C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
\end{equation}
Under these hypotheses, the symmetric hyperbolic system \ref{eqn:awe}
has a unique distribution
solution if $f \in \cal{E}'(\bR)$ (for example \cite{Lax:PDENotes},
Theorem 7.4). The Propagation of
Singularities Theorem (\cite{Tay:81}, Theorem VI.2.1) and
the pull-back property of distributions \cite{Dui:95}, Proposition 1.3.3)
imply that if $f \in \cal{E}'(\bR)$ and $(\bx_s,\bx_r) \in X_{sr}$,
then the trace $p(\bx_r, \cdot;\bx_s)$ on $\{(\bx_r,t): t \in \bR\}$ is well-defined,
as a distribution vanishing for sufficiently large
negative $t$. Define the {\em modeling operator}
\begin{equation}
  \label{eqn:modop}
  F: C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)
  \rightarrow X_{sr} \times {\cal D}'(\bR)
\end{equation}
by
\begin{equation}
  \label{eqn:modopdef}
  F[\kappa]= p|_{X_{sr} \times \bR}
\end{equation}

In this setting, acoustic FWI means: given the set of
source-receiver position pairs $X_{sr}$, source wavelet
$f \in {\cal E}'(\bR)$,  density $\rho$, and data traces
$d:  X_{sr} \rightarrow \cal{D}(\bR)$, find a bulk modulus $\kappa$
(or equivalently wave velocity $c=\sqrt{\kappa/\rho}$)
so that $F[\kappa] \approx d$ in some sense. If
$d$ and $F[\kappa]$ lie in a normed subspace of ${\cal D}'(\bR)$ for each
$(\bx_s,\bx_r) \in X_{sr}$, then the
norm provides a straightforward candidate for this sense of
approximation.

@@

\subsection{Hadamard's Decomposition}
The Hadamard decomposition of the acoustic Green's function \cite[]{Friedlander:75,Qian:JCP24} provides a
basis for a norm-based formulation of FWI. As presented by
\cite{Friedlander:75}, for instance, this construction applies to
second-order wave equations, but the system \ref{eqn:awe} is
equivalent fo a second-order problem: differentiate the first equation
with respect to $t$, eliminate $\bv$ via the second equation, and
rearrange to obtain
\begin{equation}
  \label{eqn:awe2}
  \frac{\rho}{\kappa}\frac{\partial^2 p}{\partial t^2} - \nabla^2 p -
  \nabla \log \rho \cdot \nabla p = \rho \frac{\partial f}{\partial t}
  \delta(\bx-\bx_s), \nonumber \\
  p=0, \, t \ll 0.
\end{equation}
If $(p,\bv)$ is a distribution solutions of the system
\ref{eqn:awe} then $p$ is a solution of equation
\ref{eqn:awe2}, as follows by differentiating the first equation of
\ref{eqn:awe} with respect to $t$ and eliminating $\bv$ via the second
equation. Conversely, if $p$ is a solution of the second-order initial
value problem\ref{eqn:awe2}, then under suitable conditions (including
those to be imposed below),
\[
  \bv = \beta \int_{-\infty}^t p
\]
defines a distribution, and $(p,\bv)$ solves the system \ref{eqn:awe}.

Hadamard's construction is local, in both space-time and in the set of
acoustic models. A {\em normal neighborhood} of
$\bx_s \in \bR^3$ is an open subset of $\bR^3$ 
within which every point is connected to $\bx_s$ by  a unique ray of
geometric acoustics (a geodesic for the Riemannian metric associated
to the equation \ref{eqn:awe2}) (\cite{Friedlander:75},
p. 16). Hadamard's decomposition holds in a space-time cylinder
$\Omega \times \bR$,  and $\Omega\subset \bR^3$ 
and $M \subset C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3)$ are bounded open sets satisfying 
\begin{itemize}
  \item[G1. ] $X_{sr} \subset \Omega \times \Omega$;
  \item[G2. ] for every source point $\bx_s\in X_{s}$ and $\log \kappa \in M$, $\Omega$ is a
    normal neighborhood of $\bx_s$.
  \item[G3. ] for every source point $\bx_s \in X_{s};$
    $\log \kappa \in M$, ray ${\bf X}:\bR^+\rightarrow \bR^3$
    with ${\bf X}(0)=\bx_s$ and any $t>0$, either $\{{\bf X}(s): 0\le s \le t\} \subset \Omega$
    or $\{{\bf X}(s): t < s < \infty\} \subset \bR^3 \setminus \Omega$.
  \end{itemize}
Condition G3 (the ``no return'' condition) means that once a ray
leaves $\Omega$, it does not return to $\Omega$. Conditions G1 - G3
hold for any convex open set $\Omega$ if $\kappa$ is constant, and for small perturbations and some
large perturbations of constant $\kappa$: these are open conditions in
the topology of $M$.

The arguments to follow depend an additional assumption about solutions of
the wave equation \ref{eqn:awe2}, restricted to $X_{sr}\times \bR$:
\begin{itemize}
\item[G4. ] For any $n > 0$, $f \in {\cal E}'(\bR)$, there exist $K >
  0, T \in \bR,$ and $\delta>0$ so
  that for any $k$ with $0 \le k \le n$, $(\bx_s,\bx_r) \in
 X_{sr},$
  and any $\log \kappa \in M$, the solution $p$ of \ref{eqn:awe2} is smooth
  for $t>T$ and satisfies
  \[
    |\partial_t^k p(\bx_r,t;\bx_s)| \le K e^{-\delta t}, \, t > T.
  \]
\end{itemize}
Condition G4 is a special case of {\em local energy decay}, a much
studied topic. Properties which imply G4 are in turn implied by the
{\em non-trapping} hypothesis on the model $\log \kappa \in M$: all rays of
geometric asymptotics escape to infinity (\cite{Hristova:09}, Theorem
2, \cite{EgorovShubin}, Theorem 2.104, p. 192).

Under conditions G1 - G4, the construction described by
\cite{Friedlander:75} produces for each $m \in M$ functions $\tau[\kappa]$ (travel time), $a[\kappa]$
(geometric amplitude) on $X_s \times \Omega$ and $b[\kappa]$
(remainder kernel) on $X_s \times \Omega \times \bR$, satisfying
\begin{itemize}
\item[H1. ] $\tau[\kappa], a[\kappa],$ and $b[\kappa]$ are smooth off the diagonal,
that is, $\tau[\kappa](\bx_s,\cdot) \in C^{\infty}(\Omega \setminus \{\bx_s\})$
and similarly for $a[\kappa],b[\kappa]$;
\item[H2. ] there exists $A>0$ so that for any $\log \kappa \in M$
  and $(\bx_s,\bx_r) \in X_{sr}$,
  $|\log a[\kappa](\bx_s,\bx_r)| < A$;
\item[H3. ] there exist $B \ge 0,\delta > 0$ so that for
$(\bx_s,\bx_r) \in X_{sr}$, and $\kappa \in
M$,
\begin{eqnarray}
  |b[\kappa](\bx_r,t;\bx_r,s)| &\le& B e^{-\delta t} \nonumber \\
  |\partial_t b[\kappa](\bx_r,t;\bx_r,s)| &\le& B e^{-\delta t}
  \label{eqn:best}
\end{eqnarray}
\end{itemize}
(The exponential decay estimate in item H3 follows from hypothesis G4.)

The Hadamard
decomposition of the Green's function $G[\kappa]$, i.e. the solution of the
equation \ref{eqn:awe2} for $f = \beta H$, is::
\begin{equation}
  \label{eqn:green}
  G[\kappa](\bx_s,\bx,t) = a[\kappa](\bx_s,\bx)\delta(t - \tau[\kappa](\bx_s,\bx)) +
  b[\kappa](\bx_s,\bx,t-\tau[\kappa])H(t - \tau[\kappa](\bx_s,\bx))
\end{equation}
Define $\delta_{\tau} = \delta(t-\tau)$ to be the shifted delta, that
is, $\langle \delta_{\tau},\phi\rangle = \phi(\tau)$. A useful
restatement of the definition \ref{eqn:green}, suppressing the
space-time coordinates, is
\begin{equation}
  \label{eqn:regreen}
  G[\kappa] = (a[\kappa]\delta +  (b[\kappa]H))*\delta_{\tau[\kappa]}.
\end{equation}

%Under conditions G1 - G3 above, this decomposition is the restriction of
%the Greeen's function to $\Omega \times (-\infty,t_h]$. This fact
%implies a decomposition of $F[\kappa]$, provided that the convolution $G*f$
%restricted to $[0,t_d]$ is determined by the restriction of $G$ to
%$\Omega \times [0,t_h]$. That is the case if $t_h > t_d + |\mbox{supp
%|f/2$ (the factor of $1/2$ because $f$ is odd).
%Making this assumption (which is weaker than assumptions on $t_h$ to be made below),
%\[
%   F[\kappa](\bx_s,\bx_r)(t) =
%   \phi(t)a(\bx_s,\bx_r)\rho_s\frac{df}{dt}(t-\tau[\kappa](\bx_s,\bx_r))
% \]
%\begin{equation}
%  \label{eqn:fwdgreen}
% + \phi(t)\rho_s\int
%  \,ds \, \frac{df}{dt}(t-s) b[\kappa](\bx_s,\bx_r,s-\tau[\kappa](\bx_s,\bx_r))H(s-\tau[\kappa](\bx_s,\bx_%r))
%\end{equation}

\subsection{FWI via Least Squares}
For $f \in {\cal E}'(\bR)$, the Hadamard decomposition 
(equation \ref{eqn:regreen}) gives:
\begin{equation}
  \label{eqn:fwdgreen}
  F[\kappa] = G[\kappa] * \rho \partial_t f  = \rho( a[\kappa]\delta
  +  (b[\kappa]H)) *\delta_{\tau[\kappa]}*\partial_t f.
\end{equation}
Suppose that the source wavelet $f$ satisfies
\begin{itemize}
\item[F1. ] $f \in C^{\infty}_0(\bR)$;
\item[F2. ] $f$ is odd, $f(t) = -f(-t)$.
\end{itemize}
Then equation \ref{eqn:fwdgreen}, assumptions F1, F2, bounds
\ref{eqn:best} and Young's inequality implies that $F[\kappa] \in
L^p(X_{sr} \times \bR)$ for any $p \ge 1$. Moreover there exists
$C>0$ so that for $\log \kappa\in M$, $p \ge 1$,
\begin{equation}
  \label{eqn:fwdnorm}
  \|F[\kappa]\|_{L^p(X_{sr} \times \bR)} \le C \left\|\partial_t f\right\|_{L^p(\bR)}.
\end{equation}
This observation justfies posing FWI as a least-squares problem: given $d \in
L^2(X_{sr}\times \bR)$, $f \in H^1_0(\bR)$, and $\rho_s>0$, find $\kappa$ to minimize
\begin{equation}
  \label{eqn:fwi}
  J_{\rm FWI}[m;d]= \frac{1}{2}\|F[\kappa]-d\|_{L^2(X_{sr} \times \bR)},
\end{equation}

Practical versions of least-squares FWI include various types of regularization and
imposition of constraints on $m$, see the references cited in the
Introduction.
%As mentioned there, local iterative optimization methods
%are the only practical approach to minimizing $J_{\rm FWI}$ due to the
%computational cost of approximating the forward map $F$ and related
%operators, but these tend to stagnate at highly sub-optimal
%(``cycle-skipped'') estimates of $m$ unless the initial estimate is of
%quite high quality, often available only at considerable cost in data
%collection and processing.

@@

For such extended sources,
\[
  \bar{F}[\kappa,u*w] = u*F[\kappa]
\]
Introduce the notation $S[\kappa]$ for the operator of convolution
with $F[\kappa]$, that is,
\begin{equation}
  \label{eqn:sdef}
  S[\kappa]u \equiv u*F[\kappa]
\end{equation}
Then $S[\kappa]$ is bounded on $L^2(X_{sr} \times \bR)$, uniformly in
$\log \kappa \in M$, thanks to
the bound \ref{eqn:fwdnorm} for $p=1$. $S$ also represents an
extension of $F$, in the sense that with the choice
$u(\bx_r,t;\bx_s)=\delta(t)$, the output of $F$ is recovered:
\begin{equation}
  \label{eqn:sconsist}
  S[\kappa]\delta = F[\kappa].
\end{equation}

%Since $S[\kappa]$ involves solution of the wave equation with point source
%wavelet $u*w$, application of the Hadamard decomposition \ref{eqn:green} requires a furnther
%constraint on ray geometry, expressed by condition G3. Specifically,
%\begin{itemize}
%\item[G4. ]$t_h > t_d + t_w + t_u$.
%\end{itemize}
%Then \ref{eqn:green} implies that
%\begin{equation}
%  \label{eqn:sop}
%  S[\kappa]u = w* a[\kappa] T[\kappa] ((\delta + b[\kappa]H)*u)
%\end{equation}
%in which $T[\kappa]$ is the time translation operator by $\tau[\kappa]$:
%\begin{equation}
%  \label{eqn:Tdef}
%  T[\kappa]u(\bx_r,t;\bx_s) = u(\bx_r,t-\tau[\kappa](\bx_s,\bx_r);\bx_s),
%\end{equation}
%Similarly,
%\begin{equation}
 % \label{eqn:fwdgreenop}
%  F[\kappa] = \phi a[\kappa] T[\kappa]((\delta + b[\kappa]H)*w) 
%\end{equation}

@@

\cite{Warner:16} assume (implicitly) that there is always a unique
solution of the extended inversion problem, formulated in terms of
$S$: that is, that
for any $\kappa$ and any data $d$, there is an
adaptive filter $u[\kappa;d]$ for which
$S[\kappa]u[\kappa;d]= d$. If $d$ is noise-free, that is, $d=F[\kappa]$, then the consistency relation \ref{eqn:sconsist}
suggests that $u[\kappa;d](\bx_r,t;\bx_s) = \delta(t)$. This ideal adaptive filter is in the null
space of multiplication by $t$, that is, the operator
$u(\bx_r,t;\bx_s) \mapsto t u(\bx_r,t;\bx_s)$. This
observation motivates the definition of the AWI objective function
(\cite{Warner:16}, equation XXXX)::
\begin{equation}
  \label{eqn:awideforig}
  J_{\rm AWI}[m;d] = \sum_{\bx_s,\bx_r} \left(\frac{\int dt |tu[\kappa;d](\bx_r,t;\bx_s)|^2}{\int dt |u[\kappa;d](\bx_r,t;\bx_s)|^2}\right) 
\end{equation}
in which $u[\kappa;d]$ is as defined above, that is, the solution of $S[\kappa]u[\kappa;d]
= d$. \cite{Warner:16} also formulate
AWI in terms of the inverse filter, that is, $u$ for which
$u*d = F[\kappa]$, and in fact recommend that definition. The present
discussion uses only the forward filter.
Adaptive waveform inversion consists in iterative minimization of
$J_{\rm AWI}$: choose an initial estimate $\kappa_0$ of the
model, then update $\kappa$ repeatedly using a gradient
descent method. \cite{Warner:16} describes the computation of $J_{\rm AWI}$
and its gradient in detail, and their use in a quasi-Newton
optimization algorithm.

The same observation motivates the definition of the {\em Matched
  Source Waveform Inversion} (MSWI) objective studied by
\cite{HuangSymes2015SEG,HuangSymes:Geo17}:
\begin{equation}
  \label{eqn:mswideforig}
  J_{\rm MSWI}[m;d] = \sum_{\bx_s,\bx_r} \int dt |tu[\kappa;d](\bx_r,t;\bx_s)|^2
\end{equation}
That is, the two objectives differ only by the trace-by-trace
normalization featured in the definition of AWI (equation
\ref{eqn:awideforig}). Much of the analysis of $J_{\rm AWI}$ pertains to
$J_{\rm MSWI}$, as will be explained below.
The trace-by-trace normalization in the definition of $J_{\rm AWI}$
will prove quite consequential.
